---
title: "RPM-EE Simulation Results Report"
author: "RPM-EE R analysis workflow"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
params:
  data_path: "results/export/rpm_ee_results.csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

## Methods

This report summarizes RPM-EE clinical preset simulation results. Simulation
episodes are run in Python (`src/simulation.py`'s `RPMEESimulation`), logged
per-episode with fields including `preset`, `rt_mean`, `rt_std`, and
`accuracy`, and exported to CSV/JSON by
`src/analysis/export_results_for_r.py` (see `docs/analysis_r.md` for the full
schema and export instructions). This document performs no simulation itself
-- it only reads the already-exported data below.

Episodes with `NA` `rt_mean`/`accuracy` (neutral-evidence trials, where the
DDM has no correct answer to score -- see `src/ddm.py`'s `predict_action`
docstring) are excluded from all summaries and plots below.

## Data

```{r load-data}
data_path <- params$data_path
if (!file.exists(data_path)) {
  stop(sprintf(
    "Data file not found: %s\nExport it first with:\n  python -m analysis.export_results_for_r --input <run.json> --output_csv %s",
    data_path, data_path
  ))
}

data <- read_csv(data_path, show_col_types = FALSE) %>%
  filter(!is.na(rt_mean), !is.na(accuracy))

glimpse(data)
```

## Results

### Summary statistics by preset

```{r summary-table}
summary_by_preset <- data %>%
  group_by(preset) %>%
  summarise(
    n = n(),
    rt_mean_avg = mean(rt_mean),
    rt_mean_sd = sd(rt_mean),
    accuracy_avg = mean(accuracy),
    accuracy_sd = sd(accuracy),
    .groups = "drop"
  )

knitr::kable(summary_by_preset, digits = 3)
```

## Plots

### RT distribution by preset

```{r rt-distribution, fig.width=8, fig.height=5}
ggplot(data, aes(x = rt_mean, fill = preset)) +
  geom_histogram(alpha = 0.6, position = "identity", bins = 30) +
  labs(
    title = "RT distribution by clinical preset",
    x = "Mean RT per episode (ms)", y = "Count"
  ) +
  theme_minimal()
```

### Accuracy distribution by preset

```{r accuracy-distribution, fig.width=8, fig.height=5}
ggplot(data, aes(x = preset, y = accuracy, fill = preset)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title = "Accuracy distribution by clinical preset",
    x = "Preset", y = "Accuracy"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

### Effect size vs. neurotypical baseline

Cohen's d for each non-baseline preset's `rt_mean` and `accuracy`, relative
to the `neurotypical` preset in the same dataset.

```{r calibration-plot, fig.width=8, fig.height=5}
BASELINE_PRESET <- "neurotypical"

cohens_d <- function(x, y) {
  nx <- length(x)
  ny <- length(y)
  pooled_sd <- sqrt(((nx - 1) * var(x) + (ny - 1) * var(y)) / (nx + ny - 2))
  if (pooled_sd == 0 || is.na(pooled_sd)) {
    return(NA_real_)
  }
  (mean(x) - mean(y)) / pooled_sd
}

if (BASELINE_PRESET %in% unique(data$preset)) {
  baseline_rows <- data %>% filter(preset == BASELINE_PRESET)

  effect_sizes <- data %>%
    filter(preset != BASELINE_PRESET) %>%
    group_by(preset) %>%
    summarise(
      rt_mean_d = cohens_d(rt_mean, baseline_rows$rt_mean),
      accuracy_d = cohens_d(accuracy, baseline_rows$accuracy),
      .groups = "drop"
    ) %>%
    pivot_longer(cols = c(rt_mean_d, accuracy_d), names_to = "metric", values_to = "cohens_d")

  ggplot(effect_sizes, aes(x = preset, y = cohens_d, fill = metric)) +
    geom_col(position = "dodge") +
    geom_hline(yintercept = 0, linetype = "dashed") +
    labs(
      title = sprintf("Effect size vs. %s baseline (Cohen's d)", BASELINE_PRESET),
      x = "Preset", y = "Cohen's d", fill = "Metric"
    ) +
    theme_minimal()
} else {
  message(sprintf("Baseline preset '%s' not present in data.", BASELINE_PRESET))
}
```

## Conclusions

_Fill in after reviewing the summary table and plots above: does each
clinical preset's simulated RT/accuracy differentiate from `neurotypical` in
the direction expected from the literature (see
`scripts/predictive_validation.py`'s `EXPECTED_PATTERNS` for the specific,
citation-backed expectations per preset)? Note any preset whose effect size
is near zero or in the unexpected direction for follow-up._
